Learning Discrete Bayesian Networks from Continuous Data

### 2015-12-03 BN_discretizer_paper.tex
Fix inconsistency

\begin{equation}
\label{eq:opt_prob}
\begin{aligned}
  & \sum_{i=1}^{k-1}
   - \ln
    \paren{
      1 - \exp
      \paren{
        - L \cdot  \frac{
                         \cx_{\lambda_i +1} - \cx_{\lambda_i}
                       }{
                         \cx_n - \cx_1
                        }
      }
    }
      + \sum_{i=1}^{k}
      L \cdot \frac{\cx_{\lambda_{i}} - \cx_{\lambda_{i-1} + 1}}{\cx_n - \cx_1} + \\
  & \sum_{i=1}^{k} \brock{
  \ln{{\gamma_i + J_\text{P} - 1}\choose{J_\text{P} - 1}}
  +{ \ln \paren{
        \frac{
          {\gamma_i}!
        }{
          {n^{({P})}_{i,1}}! \; {n^{({P})}_{i,2}}! \; \cdots \; {n^{({P})}_{i,J_\text{P}}}!
        }
      }}
  } + \\
  & \sum_{j=1}^{n_c} \sum_{i=1}^k \sum_{\ell=1}^{J^{(j)}_{\bm{S}}} \brock{
  { \ln
          {{n^{(j)}_{i,\ell} + J_{C_j} - 1}\choose{J_{C_j}-1}}}
    +
    \ln \paren{ {
          \frac{
            {n^{(j)}_{i,\ell}}!
          }{
            {n^{(j)}_{i,1,\ell}!} \; {n^{(j)}_{i,2,\ell}!} \; \cdots \; {n^{(j)}_{i,J_{C_j},\ell}!}
          }
        }
        } }
\end{aligned}
\end{equation}

@@

\begin{small}
  \begin{equation}
  \label{eqn:h_function}
  \begin{aligned}
  h(u,v) &=  \ln {{\gamma_{q} + J_\text{P} - 1}\choose{J_\text{P}-1}} + \ln \left( { {{\gamma_q}!}\over{ {n^{(p)}_{q,1} !} {n^{(p)}_{q,2} !} \cdots {n^{(p)}_{q,J_p} !}} } \right) \\
  & + \sum_{j=1}^{n_c} \sum_{\ell=1}^{J^{(j)}_{\bm{S}}} \brock{
    { \ln
            {{n^{(j)}_{i,\ell} + J_{C_j} - 1}\choose{J_{C_j}-1}}}
      +
      \ln \paren{ {
            \frac{
              {n^{(j)}_{i,\ell}}!
            }{
              {n^{(j)}_{i,1,\ell}!} \; {n^{(j)}_{i,2,\ell}!} \; \cdots \; {n^{(j)}_{i,J_{C_j},\ell}!}
            }
          }
          } }
  \end{aligned}
  \end{equation}
\end{small}

@@

\subsection{Auto MPG Dataset}
\label{subsec:auto}

@@

\subsubsection{Discretization with fixed Structure}
\label{subsubsec:auto_exp1}

@@

Table~\ref{table:auto_disc_table_1} lists the discretization edges and mean log-likelihoods under \num{10}-fold cross validation of the discrete Bayesian network resulting from the two discretization methods.
The MDL method does not produce any discretization edges, assigning one continuous interval to each continuous variable, and produces the result with the lower likelihood.
The reason why MDL produces fewer discretization edges is discussed in Section~\ref{subsec:discuss_exp}.
Two examples of MDL producing comparble results to the Bayesian method are given in the Appendix.

@@

\begin{figure}[ht]
  \centering
  \input{Fig5}
  \caption{The discrete-valued Bayesian network learned from the Auto MPG dataset using the Bayesian method.}
  \label{fig:auto_graph_2}
\end{figure}

@@

\subsubsection{Discretization with fixed Structure}
\label{subsubsec:wine_exp1}

@@

\begin{figure}[ht]
  \centering
  \input{Fig10}
  \caption{
    The discretization policy for variables $E$ and $K$ on the learned network.
    The discretization edge $E = 17.9$ also appears in the policies for the fixed network.
  }
  \label{fig:wine_exp2_distr}
\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Housing
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Data 3: Housing}
\label{subsec:housing}

@@

\begin{figure}[ht]
  \centering
  \input{Fig17}
  \caption{A simple Bayesian network used to demonstrate the sensitivity of each method}
  \label{fig:exp_discuss}
\end{figure}

@@

\subsection{Auto MPG with Comparable MDL Performance}
The Bayesian network structure in Figure~\ref{fig:auto_graph_3} and the discretization results in Table~\ref{table:auto_disc_table_3} demonstrate that the MDL method does not always produce fewer discretization intervals than the Bayesian approach and that the MDL method can produce discretization policies with comparable likelihood.
The Bayesian network in Figure~\ref{fig:auto_graph_3} was obtained by running Algorithm~\ref{alg:structure_learn} fifty times on the Auto MPG dataset with a maximum of two parents per variable.
The network is structured more favorably with regards to discretization than the network in Figure~\ref{fig:auto_graph_1}, as the discretization policies were simultaneously learned with the network structure, resulting in higher data likelihood.
Many edges appear in both discretization results and are bolded in Table~\ref{table:auto_disc_table_3}.

@@

\subsection{Reproducing MDL Results from the Literature}

>>>>

\begin{equation}
\label{eq:opt_prob}
\begin{aligned}
  & \sum_{i=1}^{k-1}
   - \ln
    \paren{
      1 - \exp
      \paren{
        - L \cdot  \frac{
                         \cx_{\lambda_i +1} - \cx_{\lambda_i}
                       }{
                         \cx_n - \cx_1
                        }
      }
    }
      + \sum_{i=1}^{k}
      L \cdot \frac{\cx_{\lambda_{i}} - \cx_{\lambda_{i-1} + 1}}{\cx_n - \cx_1} + \\
  & \sum_{i=1}^{k} \brock{
  \ln{{\gamma_i + J_\text{P} - 1}\choose{J_\text{P} - 1}}
  +{ \ln \paren{
        \frac{
          {\gamma_i}!
        }{
          {n^{(\text{P})}_{i,1}}! \; {n^{(\text{P})}_{i,2}}! \; \cdots \; {n^{(\text{P})}_{i,J_\text{P}}}!
        }
      }}
  } + \\
  & \sum_{j=1}^{n_c} \sum_{i=1}^k \sum_{\ell=1}^{J^{(j)}_{\bm{S}}} \brock{
  { \ln
          {{n^{(j)}_{i,\ell} + J_{C_j} - 1}\choose{J_{C_j}-1}}}
    +
    \ln \paren{ {
          \frac{
            {n^{(j)}_{i,\ell}}!
          }{
            {n^{(j)}_{i,1,\ell}!} \; {n^{(j)}_{i,2,\ell}!} \; \cdots \; {n^{(j)}_{i,J_{C_j},\ell}!}
          }
        }
        } }
\end{aligned}
\end{equation}

@@

\begin{small}
  \begin{equation}
  \label{eqn:h_function}
  \begin{aligned}
  h(u,v) &=  \ln {{\gamma_{q} + J_\text{P} - 1}\choose{J_\text{P}-1}} + \ln \left( { {{\gamma_q}!}\over{ {n^{(\text{P})}_{q,1} !} {n^{(\text{P})}_{q,2} !} \cdots {n^{(\text{P})}_{q,J_p} !}} } \right) \\
  & + \sum_{j=1}^{n_c} \sum_{\ell=1}^{J^{(j)}_{\bm{S}}} \brock{
    { \ln
            {{n^{(j)}_{i,\ell} + J_{C_j} - 1}\choose{J_{C_j}-1}}}
      +
      \ln \paren{ {
            \frac{
              {n^{(j)}_{i,\ell}}!
            }{
              {n^{(j)}_{i,1,\ell}!} \; {n^{(j)}_{i,2,\ell}!} \; \cdots \; {n^{(j)}_{i,J_{C_j},\ell}!}
            }
          }
          } }
  \end{aligned}
  \end{equation}
\end{small}

@@

\subsection{Dataset 1: Auto MPG}
\label{subsec:auto}

@@

\subsubsection{Discretization with Fixed Structure}
\label{subsubsec:auto_exp1}

@@

Table~\ref{table:auto_disc_table_1} lists the discretization edges and mean log-likelihoods under \num{10}-fold cross validation of the discrete Bayesian network resulting from the two discretization methods.
The MDL method does not produce any discretization edges, assigning one continuous interval to each continuous variable, and produces the result with the lower likelihood.
The reason why MDL produces fewer discretization edges is discussed in Section~\ref{subsec:discuss_exp}.
Two examples that MDL produces comparble results to the Bayesian method are given in the Appendix.

@@

\begin{figure}[ht]
  \centering
  \scalebox{0.9}{
  \input{Fig5}}
  \caption{The discrete-valued Bayesian network learned from the Auto MPG dataset using the Bayesian method.}
  \label{fig:auto_graph_2}
\end{figure}

@@

\subsubsection{Discretization with Fixed Structure}
\label{subsubsec:wine_exp1}

@@

\begin{figure}[ht]
  \centering
  \input{Fig10}
  \caption{
    The discretization policy for variables $E$ and $K$ on the learned network.
    The discretization edge $E = 17.9$ also appears in the policies for the fixed network.
  }
  \label{fig:wine_exp2_distr}
\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Housing
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Dataset 3: Housing}
\label{subsec:housing}

@@

\begin{figure}[ht]
  \centering
  \input{Fig17}
  \caption{A simple Bayesian network used to demonstrate the sensitivity of each method.}
  \label{fig:exp_discuss}
\end{figure}

@@

\subsection*{Auto MPG with Comparable MDL Performance}
The Bayesian network structure in Figure~\ref{fig:auto_graph_3} and the discretization results in Table~\ref{table:auto_disc_table_3} demonstrate that the MDL method does not always produce fewer discretization intervals than the Bayesian approach and that the MDL method can produce discretization policies with comparable likelihood.
The Bayesian network in Figure~\ref{fig:auto_graph_3} was obtained by running Algorithm~\ref{alg:structure_learn} fifty times on the Auto MPG dataset with a maximum of two parents per variable.
The network is structured more favorably with regards to discretization than the network in Figure~\ref{fig:auto_graph_1}, as the discretization policies were simultaneously learned with the network structure, resulting in higher data likelihood.
Many edges appear in both discretization results and are bolded in Table~\ref{table:auto_disc_table_3}.

@@

\subsection*{Reproducing MDL Results from the Literature}
